\chapter{System Testing}

\section{Testing Methodologies}
The \textbf{SmartSec} platform was tested using a combination of unit testing, integration testing, functional testing, and end-to-end (E2E) testing methodologies. The testing process covered all six backend modules, authentication flows, REST API endpoint validation, the ML IDS model inference pipeline, threat API integrations, the risk scoring engine, frontend component rendering, and Supabase database persistence.

All REST API endpoints were tested using the auto-generated FastAPI Swagger documentation at \texttt{/docs} and via direct HTTP requests using Postman. Authentication endpoints were tested for valid registration, duplicate email detection, valid and invalid login credentials, expired JWT token handling, and Google/GitHub OAuth callback flows. The \texttt{/phishing/scan} endpoint was tested with safe URLs, known phishing URLs, IP-based URLs, excessively long URLs, empty strings, and malformed inputs. IDS endpoints were tested with normal traffic profiles, high-anomaly traffic profiles, and zero-value feature vectors.

The risk scoring engine was validated by manually triggering different event types and verifying that the score delta and level classification matched the EVENT\_WEIGHTS table. Decay calculation was validated by simulating events at known timestamps and verifying the mathematical output of the exponential decay formula. Email alert triggering was verified by confirming Resend API calls for High and Critical severity events in the application logs.

Approximately 30\% of total project effort was allocated to testing activities, including test case design, execution, and validation.


\section{Test Cases}
Test cases were designed to validate each core module of the SmartSec platform. The following table presents key test cases executed during system testing.

\begin{table}[h]
\centering
\caption{Test Case Table for SmartSec AI-Based Cyber Defense Platform}
\label{tab:testcases}
\begin{tabular}{|p{1cm}|p{3.5cm}|p{3.5cm}|p{4cm}|p{1.5cm}|}
\hline
\textbf{TC No.} & \textbf{Test Case} & \textbf{Input} & \textbf{Expected Output} & \textbf{Status} \\
\hline

TC01 & User Registration & Valid username, email, password & User created, JWT token returned, 201 response & Pass \\
\hline

TC02 & Duplicate Registration & Existing email address & 409 Conflict error & Pass \\
\hline

TC03 & User Login -- Valid & Correct email \& password & JWT token returned, risk assessed & Pass \\
\hline

TC04 & User Login -- Invalid & Wrong password & 401 Unauthorized, failure event logged & Pass \\
\hline

TC05 & Google OAuth Login & Valid Supabase OAuth token & JWT token + user created/found & Pass \\
\hline

TC06 & JWT Protected Route & No Authorization header & 401 Unauthorized error & Pass \\
\hline

TC07 & Expired Token & Expired JWT token & 401 Unauthorized, re-login required & Pass \\
\hline

TC08 & Phishing Scan -- Safe URL & \texttt{https://google.com} & Verdict: Safe, low risk score & Pass \\
\hline

TC09 & Phishing Scan -- Phishing URL & Known phishing domain & Verdict: Phishing, high risk score & Pass \\
\hline

TC10 & Phishing Scan -- Empty URL & Empty string & 422 Validation error & Pass \\
\hline

TC11 & IDS Normal Traffic & Normal feature vector & is\_anomaly: false, low score & Pass \\
\hline

TC12 & IDS Anomalous Traffic & High request\_rate, error\_rate & is\_anomaly: true, high score & Pass \\
\hline

TC13 & Risk Event -- Login Failure & Failed login attempt & Score +5, Low severity event logged & Pass \\
\hline

TC14 & Risk Event -- Phishing Detected & Phishing URL scanned & Score +22, High severity, email sent & Pass \\
\hline

TC15 & Risk Event -- Brute Force & 10+ failed attempts & Score +28, Critical severity & Pass \\
\hline

TC16 & Risk Decay Calculation & Score set, time elapsed & Score decays 8\% per simulated day & Pass \\
\hline

TC17 & Notification Creation & High severity event & Notification inserted in DB & Pass \\
\hline

TC18 & Mark Notification Read & Valid notification ID & is\_read set to true in DB & Pass \\
\hline

TC19 & Dashboard Aggregation & Authenticated user & All module stats returned & Pass \\
\hline

TC20 & Profile Update & New full\_name, bio & User record updated in Supabase & Pass \\
\hline

\end{tabular}
\end{table}
