% SmartSec AI-Based Cyber Defense Platform
% Report by Shubham Prajapati
\documentclass[12pt,a4paper,final,oneside]{report}
\usepackage{geometry}
\geometry{a4paper,total={170mm,257mm}}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphics}
\usepackage{array}
\usepackage{epstopdf}
\usepackage{setspace}
\usepackage[numbers]{natbib}
\usepackage{fancyhdr}
\usepackage{xcolor}
\usepackage{lipsum}
\usepackage{setspace}
\renewcommand{\baselinestretch}{1.5}
\usepackage[pdftex]{graphicx}
\usepackage{sidecap}
\usepackage{float}
\usepackage{placeins}
\usepackage{caption}
\setlength{\headheight}{14.5pt}
\addtolength{\topmargin}{-2.5pt}

\begin{document}

\input{cover.tex}
\input{cert.tex}
\input{declaration.tex}
\input{ack.tex}
\input{abst.tex}
 \addcontentsline{toc}{chapter}{Abstract}
\tableofcontents
  \listoftables
   \addcontentsline{toc}{chapter}{List of Tables}
\listoffigures
  \addcontentsline{toc}{chapter}{List of Figures}
\input{nomenclature.tex}
\addcontentsline{toc}{chapter}{Nomenclature}
\newpage
\pagestyle{plain}
\pagenumbering{arabic}
\onehalfspacing
\pagestyle{fancy}
\renewcommand{\footrulewidth}{0.4pt}
\fancyhf{}
\rhead{SmartSec}
\lfoot{BVIMIT}
\rfoot{Page \thepage}
\clearpage\input{chap1.tex}
\clearpage\input{chap2.tex}
\clearpage\input{chap3.tex}
\clearpage\input{chap4.tex}
\clearpage\input{chap5.tex}
\clearpage\input{chap6.tex}
\clearpage\input{chap7.tex}
\clearpage\input{chap8.tex}
\clearpage

\clearpage
\chapter*{References}
\addcontentsline{toc}{chapter}{References}
\thispagestyle{empty}

\begin{enumerate}

\item FastAPI Documentation. Sebastián Ramírez, 2024. Available at: https://fastapi.tiangolo.com

\item Supabase Documentation. Supabase Inc., 2024. Available at: https://supabase.com/docs

\item React.js Documentation. Meta Platforms, Inc., 2024. Available at: https://react.dev

\item scikit-learn: Machine Learning in Python. Pedregosa et al., Journal of Machine Learning Research 12, pp. 2825--2830, 2011.

\item VirusTotal API v3 Documentation. Google LLC, 2024. Available at: https://developers.virustotal.com/reference/overview

\item Google Safe Browsing API v4 Documentation. Google LLC, 2024. Available at: https://developers.google.com/safe-browsing

\item AbuseIPDB API Documentation. Marathon Studios Inc., 2024. Available at: https://www.abuseipdb.com/api.html

\item python-jose: JSON Web Tokens in Python. Available at: https://python-jose.readthedocs.io

\item Resend Email API Documentation. Resend Inc., 2024. Available at: https://resend.com/docs

\item Richardson, L., and Ruby, S., RESTful Web Services. O'Reilly Media, 2018.

\item Vite Documentation. Evan You, 2024. Available at: https://vitejs.dev

\item IPInfo API Documentation. IPinfo LLC, 2024. Available at: https://ipinfo.io/developers

\end{enumerate}


\end{document}
